HUME: Human UCCA-based Machine-Translation Evaluation

### 2016-03-16 paper/main.tex
www

\newcommand{\com}[1]{}
\newcommand{\secref}[1]{Section~\ref{#1}}
\newcommand{\figref}[1]{Figure~\ref{#1}}
\newcommand{\tabref}[1]{Table~\ref{#1}}
\newcommand{\XXX}[1]{{\color{red}XXX #1}} % a general todo or comment
\newcommand{\oa}[1]{\footnote{\color{red}OA: #1}}
\newcommand{\bh}[1]{\footnote{\color{blue}BH: #1}}

\def\perscite#1{\newcite{#1}}
\def\parcite#1{\cite{#1}}
\def\inparcite#1{\newcite{#1}}

@@

\title{HUME: Human UCCA-Based Evaluation of Machine Translation}

@@

  Recent interest in semantics-aware MT systems underscores the importance of
  semantic evaluation for Machine Translation (MT) systems. 
  A semantic measure, sensitive to possible semantic discrepancies
  between the translation output and the source, is likely support the construction
  and tuning of semantics-based MT, much in the same way that string-based measures,
  such as BLEU, enabled progress in shallower MT.
  We present a novel human semantic evaluation measure for MT, Human
  UCCA-based MT Evaluation (HUME), building on the UCCA semantic representation scheme.
  Making progress over previous work, the proposed measure takes into account
  a wider range of semantic phenomena and does not rely on semantic annotation
  of the MT output itself, whose interpretation is often unclear.
  We experiment with four language pairs, demonstrating HUME's broad applicability,
  as well as its reliability, reflected in its inter-annotator agreement rates.

@@

These problems were partially addressed by measures that decompose over sub-parts of the evaluated
translation sentence (henceforth, {\it translation}),
most commonly over its words or n-grams,
quantifying the overlap of its sub-parts relative a reference translation.
More recent work on MT evaluation proposed measures
that decompose over semantically defined units,
quantifying the similarity of the output and the reference in terms of
their verb argument structure; the most notable of these measures is
HMEANT \cite{lo2011structured}.
While allowing to localize semantic errors in the translation,
these measures' focus on verb argument structures misses out on many pervasive semantic phenomena,
such as inter-clause linkage and nominal argument structures, and thus only provides
a limited perspective on the semantic similarity between the output and reference.
Moreover, assigning semantic structure to MT output, especially when low in quality,
is difficult and arguably ill-defined.

We propose the Human UCCA-based MT Evaluation (HUME) metric,
a human evaluation measure that decomposes over the semantic units of the sentence.
Semantic units are defined according to the 
UCCA scheme \cite{abend2013universal}, an appealing candidate for semantic analysis,
due to its cross-linguistic applicability, support for rapid annotation, and coverage
of many fundamental semantic phenomena, such as verbal, nominal and adjectival
argument structures and the inter-relations between them.

@@

We conduct experiments with four language pairs: English-German, English-Polish,
English-Romanian and English-Czech, and report efficiency (time of annotation) and
inter-annotator agreement scores, as well as qualitative comparison of the resulting
HUME scores with crowd-sourced adequacy judgements. The results show that HUME obtains 
agreement scores which are on par with those of standard sentence-level ranking measures,
but unlike them, HUME's scores do not decrease with longer sentences.\bh{We may
  have to focus more on qualitative aspects, if we just try to sell
  the measure on IAA, then the figures are not stunning.}
The paper concludes by outlining ongoing work to construct semi- and
fully-automatic variants of HUME.

@@

Scalability is addressed by employing automatic and semi-automatic approximations of human
judgements. Most commonly, such scores decompose over the sub-parts of
the translation, and quantify how many of these sub-parts appear in a reference translation,
whose validity has been manually confirmed. This decomposition further allows system developers
to localize the errors or at least unconfirmed parts within the translation.
The most commonly used measures decompose over n-grams or even individual words, e.g., 
the BLEU \cite{Papineni:2002}, NIST \cite{Doddington:2002} and METEOR measures \cite{Banerjee:2005}.
Another common approach is to determine the similarity between the reference and translation
by applying string edit distance \cite{snover2006study}.
While these measures stimulated much progress in MT research by allowing
the evaluation of massive-scale experiments,
their decomposition over words and n-grams is better suited for localizing errors
in the output of shallow MT methods, and is less instructive in identifying the semantic
nature of the errors.

@@

Perhaps the most notable attempt at semantic MT evaluation is MEANT and
its human variant HMEANT \parcite{lo2011structured}, which quantifies the similarity between
the reference and translation in terms of the overlap in
their verbal argument structures structures and associated semantic roles.
We discuss the differences between HMEANT and HUME in \secref{sec:hmeant_comp}.

@@

\begin{tikzpicture}[semithick, x=2.4cm,y=0.8cm, >=latex, baseline=7ex,
      inner sep=.3ex, outer sep=.7ex, minimum size=1.2ex,->]
  \node (ROOT) [label={\#7}] [fill=black, circle] {}
  child {node [label={\#1}] {After} edge from parent node[left]  {$L$\hspace{1.5cm}}}
  child {node (H1) [label={\#8}] [fill=black,circle] {}
    child {node [label={\#2}] {graduation} edge from parent node[right] {$P$}}
      %child {node {to Paris} edge from parent node[right] {$A$}}
      edge from parent node[left] {$H$}
    }
    child {node (H2) [label={\#9}] [fill=black,circle] {}
      child {node  {} edge from parent[draw=none]}
      child {node  {} edge from parent[draw=none]}
      child {node [label={\#3}] (Tom) {Tom} edge from parent node[right] {$A$}}
      child {node [label={\#4}] {moved} edge from parent node[right] {$P$}}
      child {node [label={\#10}] [fill=black,circle] {}
        child {node [label={\#5}] {to} edge from parent node[right] {$R$}}
        child {node [label={\#6}] {NYC} edge from parent node[right] {$C$}}
        edge from parent node[right] {$A$}}
      %child {node {to Paris} edge from parent node[right] {$A$}}
      edge from parent node[left] {$H$}
    };
    \draw (H1) -> (Tom) node[midway, right] {$A$ \hspace{0.2cm}};
  \end{tikzpicture}
\caption{\label{fig:ucca_example}
  Sample UCCA annotation. Edge labels mark UCCA categories. \#$<$index$>$ marks the node
indices.}
\end{figure}

\paragraph{Semantic Representation.}
UCCA (Universal Conceptual Cognitive Annotation, \inparcite{abend2013universal}) is a cross-linguistically applicable, lightweight
scheme for semantic annotation. Formally, a UCCA structure is a directed acyclic graph (DAG),
whose leaves correspond to the words of the text.
The graph’s nodes, called {\sc units}, are either terminals or several elements jointly
viewed as a single entity according to some semantic or cognitive consideration. Edges bear
a category, indicating the role of the sub-unit in the relation the unit
represents. \XXX{Ondrej suggests to avoid the word 'relation' here. I usually use the
term 'relation' for what is called 'role' here, so I got confused for a while.}
UCCA
currently focuses on annotating argument structures (adjectival, nominal, verbal and others) and
relations between them. The most basic notion is the {\sc scene}, which describes a movement, an
action or a state which is persistent in time. Each Scene contains one main relation, as well
as one or more participants. For example, the sentence ``After graduating, Tom moved to NYC''
contains two Scenes, whose main relations are ``graduating'' and ``John''\bh{Don't you mean ``graduating'' and
``moved''? And where did ``John'' come from?}. The participant ``John''
is a part of both Scenes, while ``NYC'' only of the latter. Further categories account for
relations between Scenes and the internal structures of complex participants and relations.
See \figref{fig:ucca_example}.

% UCCA Motivation
UCCA is a natural candidate for defining a semantic
MT evaluation measure for a number of reasons. 
First, it defines a coarse-grained and intuitive set of categories,
and consequently can be reliably annotated by non-experts, after as little as two hours
of training \cite{marinotti2014}.
Second, UCCA is a cross-linguistically applicable scheme, seeking to represent what is shared between languages,
by building on linguistic typological theory, and specifically on ``Basic Linguistic Theory''
\cite{Dixon:10a,Dixon:10b,Dixon:12}, one of the most widely used frameworks for linguistic description.
%Indeed, UCCA's cross-linguistic applicability has been demonstrated through its
%application to annotation in English, French, German and Czech\oa{Did we do experiments on Czech?}. 
The cross-linguistic applicability of UCCA has been tested in annotations of
English, French, German and Czech.
Third, the scheme has been shown to be semantically stable
in translation: UCCA annotations of translated text usually contain the same set of relationships
\cite{sulem2015conceptual}. This is an indication that the scheme reflects
a common layer of representation that in a correct translation
is mostly shared between the translation and the source. 
We note that verbal argument structures are less stable in this sense, as a verbal
argument structure may be translated or paraphrased to a nominal or adjectival
argument structure, and are thus less suitable as a basis for a semantic measure.
For example, ``after he graduated, Tom moved to NYC''
conveys a similar meaning to ``after graduation, Tom moved to NYC''.

@@

While most broad-coverage semantic work focused on shallow semantic structures,
predominantly argument structures annotated with their semantic roles, much recent
interest has been devoted to defining more complete semantic representation schemes.
The Abstract Meaning Representation (AMR) scheme \cite{banarescu2013abstract}
shares much of UCCA's motivation for defining a more complete semantic annotation.
However, using AMR is not optimal for defining a decomposition of a sentence into semantic
units as it does not ground its semantic symbols in the text,
thus not straightwardly allowing a decomposition of the sentence into sub-units.
Moreover, AMR is more fine-grained than UCCA and consequently more difficult to annotate.\oa{should we
  add that it is less cross-linguistically stable? maybe that's an overkill..
  OB: AMR has been applied to more languages, including Chinese, so I would not
  dare saying that it is less stable.}

Other approaches have been proposed for semantic representation, which are grounded in the words
and phrases of the text, and have generally represented semantic structures as bi-lexical dependencies.
Examples include the tectogrammatical layer \parcite{sgallhp:1986} used in the Prague dependency
treebanks, e.g. the Czech-English one \parcite{hajic2012announcing}, as well
as dependencies derived from Minimal Recursion Semantics representations \parcite{oepen2006discriminant}.
Nevertheless, despite their appeal as deep and intuitive means for semantic representation,
these approaches require linguistic expertise for their annotation, due to the intricate and
fine-grained distinctions they provide, and are thus in our view less suitable for basing
an interpretable MT evaluation on. 
%\newcite{Basile:12} attempted to make some aspects of the semantic annotation more accessible,
%but to our knowledge no results of this brand have been reported.

@@

This section summarises the procedure for annotation involved in computing HUME.
We mark the source sentence as $s$ and the translation sentence as $t$. 
The procedure involves two manual steps: (1) UCCA-annotating $s$, 
(2) human judgements for each semantic unit of $s$,
defined by its UCCA annotation, of whether its meaning has been preserved in $t$.
We note that (1) is performed once for every source sentence, irrespective
of the number of its translations we wish to evaluate.

\paragraph{UCCA Annotation.}
We begin by creating UCCA annotations for the source sentence, following the
UCCA guidelines\footnote{All UCCA-related resources can be found
  here: \url{http://www.cs.huji.ac.il/~oabend/ucca.html}}.
Let $L$ be the set of possible UCCA labels.
Given a sentence $s = (w_1, \ldots,w_n)$, the UCCA annotation is a DAG $G=(V,E,\ell)$,
where $\ell: V \rightarrow L$, and where $W = \{w_1,\ldots,w_n\}$ correspond to the leaves of $G$.
For every node $v \in V$, we define its {\it yield} $yld(v)$ as $v$'s descendants in $W$.
The semantic units for a sentence $s$, induced by the annotation $G$
is the set of yields of nodes in $V$. We denote it with $Units(s)$ for brevity.
We note that the translation need not be annotated.

@@

Human annotation of the semantic units requires first deciding whether
a unit is an {\it atomic unit}, i.e., that it has no meaning-bearing
sub-units, or a {\it structural unit}. In most cases, atomic units would
correspond to individual words, but they may also correspond to opaque
multi-word expressions, i.e., expressions that do not have meaningful sub-units.
When a multi-word unit is labeled as atomic, all its sub-units' annotations are discarded
from the evaluation.

@@

Structural units have sub-units (children in the UCCA graph), which may either be
atomic or structural, and represent the relation between the sub-units.
For instance, a Scene unit 
represents the relation between the main relation, participants and secondary
relations in it\XXX{again, could we avoid the word ``relation'' here? Would
``event'' be an acceptable synonym?}. Structural units are labeled as ``Adequate'' or ``Bad'', meaning
that the relation between the sub-units went wrong.

@@

\begin{compactitem}
\item The components of the translation are ordered differently from the source,
  such that the relation between them changes. For instance, where
  the sentence ``man bites dog'' is translated into a sentence meaning ``dog bites man'',
  the semantic unit corresponding to the Scene will be marked as Bad, while
  the individual atomic units ``dog'', ``bites'' and ``man'' will be marked as Green.
\item A superfluous word or phrase has been inserted into
  the translation. For instance, if ``dog bites man'' is translated into ``dog visciously bites man''.
\item One of the sub-units is missing from the translation. For instance, if ``dog bites man'' is translated into ``dog bites''.
\end{compactitem}

\XXX{Could someone review this for Ondrej:}
Note that for non-configurational languages, the relationship between the
governing and the dependent units is often expressed using morphological
properties of the dependents instead of their ordering (this holds especially for the verb and its
modifiers, less so for components of noun phrases). Since errors in morphology
should be expressed using the traffic lights on the dependent nodes, we would expect
a higher proportion of ``Adequate'' structural units for Czech and Polish
compared to German and Romanian.

We defined the labels to reflect more adequacy,
rather than fluency judgements. Specifically, annotators are instructed to
label a unit as Adequate if its translation is understandable and preserves
the meaning of the source unit, even if its fluency is impaired.

For example, consider the sentence ``After graduation, Tom moved
to NYC'' as the source $s$, see \figref{fig:ucca_example}. The atomic units are the individual words,
nodes \#1--6 and the structural nodes are \#7--10. Assume that the translation
$t$
is ``After Tom will move to Paris, graduation''\footnote{We take $t$ to
  be in English for presentation purposes. In our experiments, $t$ is in German,
  Czech, Polish or Romanian.}.
Then ``after'', ``graduation'', ``Tom'' and ``to'' are marked Green, ``moved'' is marked
Orange (as it corresponds to ``will move'' in $t$), and ``NYC''
is marked as Red, as it was mistranslated to ``Paris''. Node \#9 is marked Adequate,
as the Scene of Tom moving is properly reflected in the translation. Node \#10, which
expresses the relation between the graduation and moving Scenes, is marked Bad,
as their relation in $t$ is reversed.

@@

%composite score (uniform, sub-sets, weighted rather not -- according to categories [or empirically according to agreement])

\subsection{Annotation Interface.}

As translations may be long, we provide assistance to the annotator.
Where the annotator is asked to judge the translation quality of a unit $u$,
the part of by highlighting the part of the translation which is likely to be
most relevant for the semantic unit in question.
This is done by highlighting the words in the translation which are aligned

using automatically derived word-to-word  in order to project
this annotation to the MT output.

looking at the translation via the word alignments, and marking
on the source structure which parts have been correctly translated.
There are two kinds of components (nodes): a word or basic semantic unit, and a structural component
which contains one or more sub-components.

In order 
, and use native speakers of the target
language to annotate this output.

Since these annotators have to judge the quality of the translation,
they also needed to be fluent in English.\bh{Somewhere we could explain that we
could do a similar annotation with monolinguals, although it entails aligning reference and MT output}

If any of these changes have occurred and this damages the meaning of the aligned translation, then we mark the 
structural node as ``Bad''. 
However if the arrangement of the child nodes is essentially correct, except that one or
more of the child nodes has themselves been translated wrongly (and so is marked
orange or red), then the structural node should be 
annotated as ``Acceptable''.

\begin{figure*}[t]
    \includegraphics[width=1\textwidth]{ucca-tree-mteval}
    \caption{UCCA Tree evaluated in comparison to an aligned ``translation'' (an
	English paraphrase).}
    \label{ucca-tree-mteval}
\end{figure*}

%TODO describe figure 2 here

There are cases where we cannot usefully compare individual words in the source to words in the target.
We often need to compare translation at the phrase level. If this is the case, then we can 
 treat the structural node as a lexical unit, and mark it with the traffic light labels. This 
means that none of the children of this node will be examined in order determine the semantic score
for this sentence. 
In Figure~\ref{ucca-tree-mteval}, we can see that the phrase ``became a passion'' has been labelled as a lexical
node and it has been evaluated as ``Orange'', or partially correct, as the translation of this phrase ``loved'' 
only partially capture the semantics of the source.

We are separating lexical and structural evaluation in order to simplify evaluation and to localise errors
to their point of origin. For this reason, if a word has been translated incorrectly,
the parent node should still be correct if the number and relationship between its children are the same
as in the source. 
In Figure~\ref{ucca-tree-mteval}, we can see that even though an important child of the first scene is 
incorrectly translated (``Golf''), and the children of are ordered differently, 
the scene itself is marked as ``Acceptable''. The translation of the scene retains the structural 
relationships between the children.
\\

\begin{figure*}[t]
    \begin{center}
    \includegraphics[width=1\textwidth]{interface-annotations}
    \caption{The HUME annotation tool, based on the UCCA tool}
    \label{mttool}
    \end{center}
\end{figure*}

>>>>

\newcommand{\com}[1]{}
\newcommand{\secref}[1]{Section \ref{#1}}
\newcommand{\figref}[1]{Figure~\ref{#1}}
\newcommand{\tabref}[1]{Table~\ref{#1}}
\newcommand{\oa}[1]{\footnote{\color{red}OA: #1}}
\newcommand{\bh}[2]{\footnote{\color{blue}BH: #1}}

@@

\title{HUME: Human UCCA-based Machine-Translation Evaluation}

@@

  Recent interest in semantics-aware MT systems underscores the importance of
  semantic evaluation for Machine Translation (MT) systems. 
  A semantic measure, sensitive to possible semantic discrepencies
  between the translation output and the source, is likely support the construction
  and tuning of semantics-based MT, much in the same way that string-based measures,
  such as BLEU, enabled progress in shallower MT.
  We present a novel human semantic evaluation measure for MT, Human
  UCCA-based MT evaluation (HUME), building on the UCCA semantic representation scheme.
  Making progress over previous work, the proposed measure takes into account
  a wider range of semantic phenomena and does not rely on semantic annotation
  of the MT output itself, whose interpretation is often unclear.
  We experiment with four language pairs, demonstrating HUME's broad applicability,
  as well as its reliability, reflected in its inter-annotator agreement rates.

@@

These problems were partially addressed by measures that decompose over sub-parts of the evaluated
translation sentence (henceforth, {\it translation}),
most commonly over its words or n-grams,
quantifying the overlap of its sub-parts relative a reference translation.
More recent work on MT evaluation proposed measures
that decompose over semantically defined units,
quantifying the similarity of the output and the reference in terms of
their verb argument structure (the most notable of these measures is
HMEANT \cite{lo2011structured}).
While allowing to localize semantic errors in the translation,
these measures' focus on verb argument structures misses out on many pervasive semantic phenomena,
such as inter-clause linkage and nominal argument structures, and thus only provides
a limited perspective on the semantic similarity between the output and reference.
Moreover, assigning semantic structure to MT output, especially when low in quality,
is difficult and arguably ill-defined.

We propose the Human UCCA-based MT Evaluation ({\it HUME}) metric,
a human evaluation measure that decomposes over the semantic units of the sentence.
Semantic units are defined according to the 
UCCA scheme \cite{abend2013universal}, an appealing candidate for semantic analysis,
due to its cross-linguistic applicability, support for rapid annotation, and coverage
of many fundamental semantic phenomena, such as verbal, nominal and adjectival
argument structures and the inter-relations between them.

@@

We conduct experiments with four language pairs: English-German, English-Polish,
English-Romanian and English-Czech, and report efficiency (time of annotation) and
inter-annotator agreement scores, as well as qualitative comparison of the resulting
HUME scores with crowd-sourced adequacy judgements. Results show that HUME obtains 
agreement scores which are on par with those of standard sentence-level ranking measures,
but unlike them, HUME's scores do not decrease with longer sentences.\bh{We may
  have to focus more on qualitative aspects, if we just try to sell
  the measure on IAA, then the figures are not stunning.}
The paper concludes by outlining ongoing work to construct semi- and
fully-automatic variants of HUME.

@@

Scalability is addressed by employing automatic and semi-automatic approximations of human
judgements. Most commonly, such scores decompose over the sub-parts of
the translation, and quantify how many of these sub-parts appear in a reference translation,
whose validity has been manually confirmed. This decomposition further allows system developers
to localize the errors within the translation.
The most commonly used measures decompose over n-grams or even individual words, e.g., 
the BLEU \cite{Papineni:2002}, NIST \cite{Doddington:2002} and METEOR measures \cite{Banerjee:2005}.
Another common approach is to determine the similarity between the reference and translation
by applying string edit distance \cite{snover2006study}.
While these measures stimulated much progress in MT research by allowing
the evaluation of massive-scale experiments,
their decomposition over words and n-grams is better suited for localizing errors
in the output of shallow MT methods, and is less instructive in identifying the semantic
nature of the errors.

@@

Perhaps the most notable attempt at semantic MT evaluation is MEANT and
its human variant HMEANT, which quantifies the similarity between
the reference and translation in terms of the overlap in
their verbal argument structures structures, and associated semantic roles.
We discuss the differences between HMEANT and HUME in \secref{sec:hmeant_comp}.

@@

\begin{tikzpicture}[semithick, x=2.4cm,y=0.8cm, >=latex, baseline=7ex,
      inner sep=.3ex, outer sep=.7ex, minimum size=1.2ex,->]
  \node (ROOT) [label={\#7}] [fill=black, circle] {}
  child {node [label={\#1}] {After} edge from parent node[left]  {$L$\hspace{1.5cm}}}
  child {node (H1) [label={\#8}] [fill=black,circle] {}
    child {node [label={\#2}] {graduation} edge from parent node[right] {$P$}}
      %child {node {to Paris} edge from parent node[right] {$A$}}
      edge from parent node[left] {$H$}
    }
    child {node (H2) [label={\#9}] [fill=black,circle] {}
      child {node  {} edge from parent[draw=none]}
      child {node  {} edge from parent[draw=none]}
      child {node [label={\#3}] (Tom) {Tom} edge from parent node[right] {$A$}}
      child {node [label={\#4}] {moved} edge from parent node[right] {$P$}}
      child {node [label={\#10}] [fill=black,circle] {}
        child {node [label={\#5}] {to} edge from parent node[right] {$R$}}
        child {node [label={\#6}] {NYC} edge from parent node[right] {$C$}}
        edge from parent node[right] {$A$}}
      %child {node {to Paris} edge from parent node[right] {$A$}}
      edge from parent node[left] {$H$}
    };
    \draw (H1) -> (Tom) node[midway, right] {$A$ \hspace{0.2cm}};
  \end{tikzpicture}
\caption{\label{fig:ucca_example}
  Sample UCCA annotation. Edge labels mark UCCA categories. \#<index> marks the node
indices.}
\end{figure}

\paragraph{Semantic Representation.}
UCCA (Universal Conceptual Cognitive Annotation) is a cross-linguistically applicable, lightweight
scheme for semantic annotation. Formally, a UCCA structure is a directed acyclic graph (DAG),
whose leaves correspond to the words of the text.
The graph’s nodes, called {\sc units}, are either terminals or several elements jointly
viewed as a single entity according to some semantic or cognitive consideration. Edges bear
a category, indicating the role of the sub-unit in the relation the unit represents. UCCA
currently focuses on annotating argument structures (adjectival, nominal, verbal and others) and
relations between them. The most basic notion is the {\sc scene}, which describes a movement, an
action or a state which is persistent in time. Each Scene contains one main relation, as well
as one or more participants. For example, the sentence ``After graduating, Tom moved to NYC''
contains two Scenes, whose main relations are ``graduating'' and ``John''\bh{Don't you mean ``graduating'' and
``moved''? And where did ``John'' come from?}. The participant ``John''
is a part of both Scenes, while ``NYC'' only of the latter. Further categories account for
relations between Scenes and the internal structures of complex participants and relations.
See \figref{fig:ucca_example}.

% UCCA Motivation
UCCA is a natural candidate for defining a semantic
MT evaluation measure for a number of reasons. 
First, it defines a coarse-grained and intuitive set of categories,
and consequently can be reliably annotated by non-experts, after as little as two hours
of training \cite{marinotti2014}.
Second, UCCA is a cross-linguistically applicable scheme, seeking to represent what is shared between languages,
by building on linguistic typological theory, and specifically on ``Basic Linguistic Theory''
\cite{Dixon:10a,Dixon:10b,Dixon:12}, one of the most widely used frameworks for linguistic description.
Indeed, UCCA's cross-linguistic applicability has been demonstrated through its
application to annotation in English, French, German and Czech\oa{Did we do experiments on Czech?}. 
Third, the scheme has been shown to be semantically stable
in translation: UCCA annotations of translated text usually contain the same set of relationships
\cite{sulem2015conceptual}. This is an indication that the scheme reflects
a common layer of representation that in a correct translation
is mostly shared between the translation and the source. 
We note that that verbal argument structures are less stable in this sense, as a verbal
argument structure may be translated or paraphrased to a nominal or adjectival
argument structure, and are thus less suitable as a basis for a semantic measure.
For example, ``after he graduated, Tom moved to NYC''
conveys a similar meaning to ``after graduation, Tom moved to NYC''.

@@

While most broad-coverage semantic work focused on shallow semantic structures,
predominantly argument structures annotated with their semantic roles, much recent
interest has been devoted to defining more complete semantic representation schemes.
The Abstract Meaning Representation (AMR) scheme \cite{banarescu2013abstract}
shares much of UCCA's motivation for defining a more complete semantic annotation.
However, using AMR is not optimal for defining a decomposition of a sentence into semantic
units as it does not ground its semantic symbols in the text,
thus not straightwardly allowing a decomposition of the sentence into sub-units.
Moreover, AMR is more fine-grained than UCCA and consequently more difficult to annotate.\oa{should we
  add that it is less cross-linguistically stable? maybe that's an overkill..}

Other approaches have been proposed for semantic representation, which are grounded in the words
and phrases of the text, and have generally represented semantic structures as bi-lexical dependencies.
Examples include the Prague Treebank tectogrammatical layer \cite{hajic2012announcing}, as well
as dependencies derived from Minimal Recursion Semantics representations \cite{oepen2006discriminant}.
Nevertheless, despite their appeal as deep and intuitive means for semantic representation,
these approaches require linguistic expertise for their annotatation, due to the intricate and
fine-grained distinctions they provide, and are thus in our view less suitable for basing
an interpretable MT evaluation on. 
%\newcite{Basile:12} attempted to make some aspects of the semantic annotation more accessible,
%but to our knowledge no results of this brand have been reported.

@@

This section summarises the manual annotation procedure performed in order
to compute the HUME measure. 
We mark the source sentence as $s$ and the translation sentence as $t$. 
The procedure involves two manual steps: (1) UCCA-annotating $s$, 
(2) human judgements as to the translation quality of each semantic unit of $s$,
defined by its UCCA annotation, relative to $t$.
We note that UCCA annotation is performed once for every source sentence, irrespective
of the number of its translations we wish to evaluate.

\paragraph{UCCA Annotation.}
We begin by creating UCCA annotations for the source sentence, following the
UCCA guidelines\footnote{All UCCA-related resources can be found
  here: http://www.cs.huji.ac.il/~oabend/ucca.html}.
Let $L$ be the set of possible UCCA labels.
Given a sentence $s = (w_1, \ldots,w_n)$, its UCCA annotation is a DAG $G=(V,E,\ell)$,
where $\ell: V \rightarrow L$, and where $W = \{w_1,\ldots,w_n\}$ correspond to the leaves of $G$.
For every node $v \in V$, we define its {\it yield} $yld(v)$ as $v$'s descendants in $W$.
The semantic units for a sentence $s$ induced by the annotation $G$,
is the set of yields of nodes in $V$. %We denote it with $Units(s)$ for brevity.
We note that the translation is not UCCA- annotated.

\begin{figure*}[t]
    \includegraphics[width=1\textwidth]{ucca-tree-mteval}
    \caption{UCCA Tree evaluated in comparison to an aligned ``translation'' (an
	English paraphrase).}
    \label{ucca-tree-hume}
\end{figure*}

@@

Human annotation of the semantic units requires first deciding whether
a unit is {\it atomic}, i.e., that it has no meaning-bearing
sub-units, or {\it structural}. In most cases, atomic units 
correspond to individual words, but they may also correspond to unanalyzable
multi-word expressions, i.e., expressions that do not have meaningful sub-units.
When a multi-word unit is labeled as atomic, all its sub-units are discarded
from the evaluation.

@@

Structural units have sub-units (children in the UCCA graph), which may either be
atomic or structural, and represent the relation between the sub-units.
For instance, a Scene unit 
represents the relation between the main relation, participants and secondary
relations in it. Structural units may be labeld as ``Adequate'' or ``Bad'', meaning
that the relation between the sub-units went wrong.

@@

\begin{compactitem}
\item The components of the translation are ordered differently from the source,
  such that the relation between them changes. For instance, where
  the sentence ``man bites dog'' is translated into a sentence meaning ``dog bites man'',
  the semantic unit corresponding to the Scene will be marked as Bad, while
  the inidividual atomic units ``dog'', ``bites'' and ``man'' will be marked as Green.
\item A superfluous word or phrase has been inserted into
  the translation. For instance, if ``dog bites man'' is translated into ``dog visciously bites man''.
\item One of the sub-units is missing from the translation. For instance, if ``dog bites man'' is translated into ``dog bites''.
\end{compactitem}

We note that the defined labels are more similar in spirit to adequacy,
rather than fluency judgements. Specifically, annotators are instructed to
label a unit as Adequate if its translation is understandable and preserves
the meaning of the source unit, even if its fluency is impaired.

For example, consider the sentence ``After graduation, Tom moved
to NYC'' \figref{fig:ucca_example}. The atomic units are the inidividual words,
nodes \#1--6 and the structural nodes are \#7--10. Assume that the translation
is ``After Tom will move to Paris, graduation''\footnote{We take $t$ to
  be in English for presentation purposes. In our experiments $t$ is in German,
  Czech, Polish or Romanian.}.
Then ``after'', ``graduation'', ``Tom'' and ``to'' are marked Green, ``moved'' is marked
Orange (as it corresponds to ``will move'' in $t$), and ``NYC''
is marked as Red, as it was mistranslated to ``Paris''. Node \#9 is marked Adequate,
as the Scene of Tom moving is properly reflected in the translation. Node \#10, which
expresses the relation between the graduation and moving Scenes, is marked Bad,
as their relation in $t$ is reversed.

Another example is given in \figref{fig:ucca-tree-hume}.
%``Golf became an obsession for his oldest daughter, she took daily lessons and became very good'',
The figure shows an annotation of a UCCA tree, and an example annotation for
each of the units\oa{annotation for ``she'' is missing. and for ``very good'' is wrong}.
We see that ``became a passion'' is deemed atomic, as it is translated as a single unit
in $t$, and thus its sub-units, ``became'', ``a'' and ``passion'' are not evaluated
individually.\oa{I'm not sure I agree with the UCCA annotation}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Annotation Interface}

We turn to discuss the interface for collecting human judgements on the semantic units of the source,
relating to the translation. UCCA annotations of the source are done offline, using the UCCA web
application\footnote{See UCCA webpage.}.

\figref{fig:interface} presents the user interface of HUME. The sentence at the top
shows the complete MT system output. The source sentence appears at the top, and
the source translation directly below it. Underneath the source sentence appears
its list of semantic units. For each semantic unit, including the one for the
entire source sentence, there is a textbox, displaying the yield of that unit.
Information on the category of the unit (e.g., whether it is a Scene, participant etc.).
The user is asked to select a label for each unit,
by clicking the ``A'', ``B'', Green, Orange, or Red buttons to the left of the unit's box.

The interface presents, for each unit, the segment of the translation which aligns to it.
This allows the user, especially in long sentences, to focus her attention on the parts
mostly likely to be relevant for her judgement. As the alignments are automatically derived,
and therefore noisy, the annotator is instructed to treat the aligned text is a cue, but to ignore
it and make a judgement according to the full translation sentence, in cases where the alignment
is misleading.

Concretely, let $s = (w_1,\ldots,w_n)$ be a source sentence, $t = (w'_1,\ldots,w'_n)$ a translation,
and $A \subset 2^{s \times t}$ a many-to-many word alignment between them.
If $u$ is a semantic unit in $s$, whose yield is $yld(u)$, we define the aligned text in
$t$ to be $a(u)=\bigcup_{(x_s,x_t) \in A \wedge x_s \cap yld(u) \neq \emptyset} x_t \subset t$.
In the event that $a(u)$ is discontiguous in the text, we mark words between the left
and right boundaries of $a(u)$ which are not contained in it (intervening words)
in a smaller red font. Intervening words are likely change the meaning of the translation
of $u$, and thus should be paid attention to when considering whether the translation
is correct or not.

For example, consider the sentence $s=$``The man watched the dog'' and the German
translation ``t=Der Mann sieht den Hund an''. Then ``watched'' will be aligned to
``sieht ... an'', and ``den Hund'' will be marked as intervening word. Another example
can be seen in \figref{fig:interface}, where ``ongoing pregnancy'' is translated to
``Schwangerschaft ... laufenden'' (lit. ``pregnancy ... ongoing'').

\

%There are cases where we cannot usefully compare individual words in the source to words in the target.
%We often need to compare translation at the phrase level. If this is the case, then we can 
% treat the structural node as a lexical unit, and mark it with the traffic light labels. This 
%means that none of the children of this node will be examined in order determine the semantic score
%for this sentence. 
%In Figure~\ref{ucca-tree-mteval}, we can see that the phrase ``became a passion'' has been labelled as a lexical
%node and it has been evaluated as ``Orange'', or partially correct, as the translation of this phrase ``loved'' 
%only partially capture the semantics of the source.

\begin{figure*}[t]
    \begin{center}
    \includegraphics[width=1\textwidth]{hume_interface.png}
    \caption{The HUME annotation tool, based on the UCCA tool. The faded orange box at the top
      presents the translation, the large box below it presents the source. The tree of semantic units
      for the source is given underneath, where the aligned text appears in italics, and intervening words
      appear in red with a smaller font (e.g., in the first box).}
    \label{fig:interface}
    \end{center}
\end{figure*}
